MLH1 Data Report

### 2019-09-26 doc/Main_Report.rmd
re-ran main report with new batch 16 data

#load main data file
load(file="data/MLH1/MLH1_data_setup_8.29.19.RData")

@@

```{r quant.repeatability, echo=FALSE, include=TRUE}
#load, clean up and mat one of the master.MLH1count sheets
#master
old.master.MLH1 = read.csv("~./MLH1repo/data/old_master_MLH1_nov15.csv", header = TRUE)
#remove all with 'X'
#create a real fileName col, add rev.tif,

@@

uutesting <- add_fullFileName(old.master.MLH1)

@@

#match old.full.filname with MLH1$fileName

@@

#the true false might have to be changed first

chrm.class.DF$same_count_num <- as.numeric(chrm.class.DF$same_count)

#regression of quality and MLH1.dif
lmreg <- lm(abs(MLH1.dif) ~ quality + nMLH1.foci, data= chrm.class.DF)
#summary(lmreg)
#according to this linear regression, quality is not a significant

#make regression for just same counts
#I tried making a  binomial regression model, 
#but quality is bad predictor, only 5 levels)
lmreg2 <- glm(same_count ~ quality, data= chrm.class.DF, family=binomial(link="logit"))
#summary(lmreg2)
#to plot the model / bimodal data
#fit = glm(same_count_num ~ as.numeric(quality), data=chrm.class.DF, family=binomial)
#newdat <- data.frame(quality=seq(min(chrm.class.DF$quality), max(chrm.class.DF$quality),len=102))
#newdat$vs = predict(fit, data=newdat, type="response")
#the prediction is not going 1 or 0 -- so doesn't fit logistic
#plot(vs ~ quality, data=newdat, col="red4")
#lines(vs ~ quality, newdat, col="green4", lwd=2)

>>>>

#load main data file
load(file="data/MLH1/MLH1_data_setup_9.26.19.RData")#added batch16
#load(file="data/MLH1/MLH1_data_setup_8.29.19.RData") #

@@

In the original_DF and MLH1 DF's there are ~400 and X respectively of cells (filenames) that are duplicated, or cells that have been counted in more than 1 quant batch

Datasets / sources of repeat measures
-chrom proportions
-multiple batch counts (later than batch0) *focus on this dataset for now. since its >100 and fairly easy to compile
-old and new (batch0)

The main take away is that higher quality scored cells have lower error in MLH1 counts.

> Does the amount of error differ across strains or sexes? (female might have a higher error rate)

> What sort of adjustments or tests should be made to results or analyses given the known error rate?

```{r quant.repeatability_rep1, echo=FALSE, include=TRUE}
#set up data for combining replicate counts

#1. replicates in DF, (17 mice, across 10 batches), focus on those from original for now

#isolate the duplicates (fileName) from main MLH1 DFs
n_occur <- data.frame(table(original_DF$fileName))
n_occur[n_occur$Freq > 1,]
original_DF_duplicates <- original_DF[original_DF$fileName %in% n_occur$Var1[n_occur$Freq > 1],]#2 is the max freq of duplicates,

#are there any batch, fileName duplicates?
original_DF_duplicates$unique.key <-  paste(original_DF_duplicates$fileName, original_DF_duplicates$Batch, sep = "_")
anyDuplicated(original_DF_duplicates$unique.key)#31 duplicated fileName_batch ... really 2?

n_occur <- data.frame(table(original_DF_duplicates$unique.key))
n_occur[n_occur$Freq > 1,]
double.trouble <- original_DF_duplicates[original_DF_duplicates$unique.key %in% n_occur$Var1[n_occur$Freq > 1],]
#2 fileNames from batch 0 appear 2x, remove them from DF

original_DF_duplicates <- original_DF_duplicates[ ! original_DF_duplicates$fileName %in% double.trouble$fileName, ]#removes 4 rows

#seperated duplicates into categories out in Excel (write to /results, moved final file to /data)
#This data frame was made ... with Jered's sugg
#write the duplicate file to csv/excel
#sort by file name
#need to remove '?' -- deal with the ? and X's
#write.table(original_DF_duplicates, "~./MLH1repo/results/orginal_DF_duplicates.csv", sep=",", row.names = FALSE)

original_DF_duplicates.notes = read.csv("~./MLH1repo/data/orginal_DF_duplicates.csv", header = TRUE)
#copies of fileNames can be distinguished with the 'separate.DF' column

#remove '?' that I stupidly left in
original_DF_duplicates.notes$nMLH1.foci = gsub("\\?", "", original_DF_duplicates.notes$nMLH1.foci)
original_DF_duplicates.notes$quality = gsub("\\?", "", original_DF_duplicates.notes$quality)
#these lines change col to chr

original_DF_duplicates.notes <- droplevels(original_DF_duplicates.notes)
#now there should be no more '?' going forward

#original_DF_duplicates.notes$quality <- as.factor(original_DF_duplicates.notes$quality)

#make lists of rows with "" or X's so I can remove all fileNames for those
#these lists just have na
quality.na.list <- original_DF_duplicates.notes$fileName[which(is.na(original_DF_duplicates.notes$quality) | original_DF_duplicates.notes$quality=="")] #16

MLH1.na.list <- original_DF_duplicates.notes$fileName[which(is.na(original_DF_duplicates.notes$nMLH1.foci) | original_DF_duplicates.notes$nMLH1.foci=="") ]#22

X.list <- original_DF_duplicates.notes$fileName[grepl("X", original_DF_duplicates.notes$X)]#40

#make DF which only has 
matches.only <- original_DF_duplicates.notes[ ! original_DF_duplicates.notes$fileName %in% quality.na.list, ]
matches.only <- matches.only[ ! matches.only$fileName %in% MLH1.na.list, ]
matches.only <- matches.only[ ! matches.only$fileName %in% X.list, ]

#str(matches.only)
#clean up match list (chr to num)
matches.only$nMLH1.foci <- as.numeric(matches.only$nMLH1.foci)
matches.only$quality <- as.numeric(matches.only$quality)

levels(matches.only$nMLH1.foci) #leves don't appear in num?.. move on, dwabt

#check that the 1, and 2 are equal number
length(matches.only$Original.Name[matches.only$separate.DF == 1]) #both 178

#i want to measure correleation of MLH1 across bins of quality (1,2,3,4,5)
#make new df with filename and MLH1.1 and mlh1.2

dups.here <- data.frame(  FN = unique(matches.only$Original.Name), MLH1.1 = matches.only$nMLH1.foci[matches.only$separate.DF == 1], MLH1.2 = matches.only$nMLH1.foci[matches.only$separate.DF == 2],
                        qual.1 = matches.only$quality[matches.only$separate.DF == 1] , 
                        qual.2 = matches.only$quality[matches.only$separate.DF == 2]
                          ) #OMG this worked!!!
dups.here$diff <- abs(dups.here$MLH1.1 - dups.here$MLH1.2) #the aveage difference is pretty small, but some have very high difference
#the high differences could potentially be mix ups

repeatability.table <- ddply(.data=dups.here, 
                 .(qual.1, qual.2),
                 summarize, 
                 ncells = length(unique(FN)),
                 plain_cor_MLH1 = cor(MLH1.1,MLH1.2)
)
#this table shows how many cells went changed into different qual score bins
#easiest patter, 1-1, 1-2 (high correlation in MLH1)
#5-5 lowest, (5-3 also low)
#4-5 has high corr? (direction shouldn't matter)
#4-2 has negative corr?

#of the different quality permutation bins, 37% don't change quality bin, 44% of cells change by 1 (either direction)

no.change = c(14,13,17,13,10)#67
change.1 = c(10,7,8,20,9,14,4,7)#79

repeatability.table.qual1 <- ddply(.data=dups.here, 
                 .(qual.1),
                 summarize, 
                 ncells = length(unique(FN)),
                 plain_cor_MLH1 = cor(MLH1.1,MLH1.2),
                 dif = mean(MLH1.1 - MLH1.2)
)

repeatability.table.qual2 <- ddply(.data=dups.here, 
                 .(qual.2),
                 summarize, 
                 ncells = length(unique(FN)),
                 plain_cor_MLH1 = cor(MLH1.1,MLH1.2)
)
#These tables show that qual1 cells have the highest corrleation of MLH1
dups.here.plain_cor_qual <- cor(dups.here$qual.1, dups.here$qual.2)

#i don't know how to compare quality in more

#nMLHJ1, quality are factors
original_DF_duplicates.notes$quality2 = gsub("\\?", "", original_DF_duplicates.notes$quality)

original_DF_duplicates.notes <- droplevels(original_DF_duplicates.notes)

#factor > chr > number  (this col is now chr is can go to num)
#original_DF_duplicates.notes$MLH1.foci2 <- as.numeric(original_DF_duplicates.notes$MLH1.foci2)
#original_DF_duplicates.notes$quality2 <- as.numeric(original_DF_duplicates.notes$quality2)

#one <- original_DF_duplicates.notes[original_DF_duplicates.notes$separate.DF == 1,]
#two <- original_DF_duplicates.notes[original_DF_duplicates.notes$separate.DF == 2,]
#rename this
#plain_cor_MLH1 <- cor(one$MLH1.foci2,two$MLH1.foci2,use = "complete.obs")
#.857
#plain_cor_qual <- cor(one$quality2,two$quality2,use = "complete.obs")
#.63

#What's the relationship of MLH1 cor with quality?
#what the rpeat of quality score?
#what are the correlations across quality score bins?

#issues that complicate the plain calculation
#some duplicates were 'x'ed and some weren't -- this causes different numbers
#connected with above, some duplicates have mlh1 scores, other don't -- this kinda messes up correlation

original_DF_duplicates.notes <- original_DF_duplicates.notes[ !grepl("X", original_DF_duplicates.notes$X) , ]
original_DF_duplicates.notes <- original_DF_duplicates.notes[!(is.na(original_DF_duplicates.notes$nMLH1.foci) | original_DF_duplicates.notes$nMLH1.foci==""), ]
#this causes unequal lengths

original_DF_duplicates.notes <- droplevels(original_DF_duplicates.notes)
original_DF_duplicates.notes$nMLH1.foci <- as.character(original_DF_duplicates.notes$nMLH1.foci)

original_DF_duplicates.notes$nMLH1.foci <- as.numeric(original_DF_duplicates.notes$nMLH1.foci)

#correlationsnMLH1 
#one <- original_DF_duplicates.notes[original_DF_duplicates.notes$separate.DF == 1,]
#two <- original_DF_duplicates.notes[original_DF_duplicates.notes$separate.DF == 2,]
#plain_cor <- cor(one$nMLH1.foci,two$nMLH1.foci)
#old.master.MLH1 <- original_DF[ ! original_DF$fileName %in% ll$fileName, ]

#check that all batches are distinguishable across duplicate fileNames
#make unique key fileName_batch -- then check for duplicate of that 
#original_DF$Batch <- factor(original_DF$Batch, ordered = TRUE, levels = c("Batch0", "Batch1", "Batch2", "Batch3", "Batch4","Batch5","Batch6",  "Batch7", "Batch8","Batch9","Batch10", "Batch11","Batch12", "Batch13", "Batch14","Batch15"))
#after doing this, original_DF$Batch[2000] > original_DF$Batch[10] will be evaluated!

```{r repeatability_rep2, echo=FALSE, include=FALSE, eval=FALSE}
#2. old vs new, (remove duplicates and merge batch0 with old_master_mlh1)
#load, clean up and mat one of the master.MLH1count sheets
#master
old.master.MLH1 = read.csv("~./MLH1repo/data/old_master_MLH1_nov15.csv", header = TRUE)
#remove all with 'X'
#create a real fileName col, add rev.tif,

@@

old.master.MLH1 <- add_fullFileName(old.master.MLH1)
old.master.MLH1$fileName <- old.master.MLH1$file.name_full

@@

#need to test for duplicates before mergeing?
n_occur <- data.frame(table(old.master.MLH1$file.name_full))
n_occur[n_occur$Freq > 1,]
ll <- old.master.MLH1[old.master.MLH1$file.name_full %in% n_occur$Var1[n_occur$Freq > 1],]
old.master.MLH1 <- old.master.MLH1[ ! old.master.MLH1$fileName %in% ll$fileName, ]

n_occur <- data.frame(table(original_DF$fileName))
n_occur[n_occur$Freq > 1,]
original_DF_duplicates <- original_DF[original_DF$fileName %in% n_occur$Var1[n_occur$Freq > 1],]
table(original_DF_duplicates$mouse)#17 mice

old.master.MLH1 <- original_DF[ ! original_DF$fileName %in% ll$fileName, ]

#old.master.MLH1 <- original_DF[ ! original_DF$fileName %in% ll$fileName, ]

#there are duplicates in original_DF and MLH1

uutesting2 <- merge(old.master.MLH1, original_DF, by.x = fileName, by.y = Original.Name)
#match old.full.filname with MLH1$fileName
```

```{r repeatability_rep3, echo=FALSE, include=FALSE}
chrm.class.DF = read.csv("~./MLH1repo/data/ChrmClassProportions.csv", header = TRUE)

@@

#the true false might have to be changed first
#chrm.class.DF$same_count_num <- as.numeric(chrm.class.DF$same_count)

#regression of quality and MLH1.dif
#lmreg <- lm(abs(MLH1.dif) ~ quality + nMLH1.foci, data= chrm.class.DF)
#summary(lmreg)
#according to this linear regression, quality is not a significant

#make regression for just same counts
#I tried making a  binomial regression model, 
#but quality is bad predictor, only 5 levels)
#lmreg2 <- glm(same_count ~ quality, data= chrm.class.DF, family=binomial(link="logit"))
#summary(lmreg2)
#to plot the model / bimodal data
#fit = glm(same_count_num ~ as.numeric(quality), data=chrm.class.DF, family=binomial)
#newdat <- data.frame(quality=seq(min(chrm.class.DF$quality), max(chrm.class.DF$quality),len=102))
#newdat$vs = predict(fit, data=newdat, type="response")
#the prediction is not going 1 or 0 -- so doesn't fit logistic
#plot(vs ~ quality, data=newdat, col="red4")
#lines(vs ~ quality, newdat, col="green4", lwd=2)
